Outbreak Threshold Detection

### 2023-03-01 notebooks/Julia/simulation-detection.qmd
Update initial test SIR to add births and deaths Needed to switch to specifying the rates and affects directly as MTK implementation wouldn't solve with tau-leaping. Might be worth re-visiting as would be useful to use tau-leaping when running multiple simulations.
Updated code implementation to explicitly use frequency-dependent
transmission model (note `rate1` includes `/N` term).

```{julia}
using JumpProcesses, StochasticDiffEq, Random, Statistics, LinearAlgebra
using DataFrames, CairoMakie, ProgressMeter

@@

```{julia}
β = 0.1 / 1000.0
ν = 0.01;
p = (β, ν)
```

```{julia}
tmax = 250.0
δt = 0.001
tspan = (0.0, tmax)
tlength = length(tspan[1]:δt:tspan[2])
Random.seed!(1234);
```

```{julia}
function rate(out, u, p, t)
    out[1] = p[1] * u[1] * u[2]   # β * S * I
    out[2] = p[2] * u[2]          # ν * I
    nothing
end
```

```{julia}
c = zeros(3, 2)
# S + I --> I
c[1, 1] = -1    # S -> S - 1
c[2, 1] = 1     # I -> I + 1

# I --> R
c[2, 2] = -1    # I -> I - 1
c[3, 2] = 1     # R -> R + 1
```

```{julia}
function change(du, u, p, t, counts, mark)
    mul!(du, c, counts)
    nothing
end
```

```{julia}
rj = RegularJump(rate, change, 2)

u₀ = [1000.0, 50.0, 0.0]
prob = DiscreteProblem(u₀, tspan, p)
jump_prob = JumpProblem(prob, Direct(), rj)

sol = solve(jump_prob, TauLeaping(); dt = δt)
```

```{julia}
sir_array = zeros(Float64, 4, tlength) 
sir_array[1:3, :] = Array(sol)
sir_array[4, :] = sol.t
```

```{julia}
f = Figure()
ax = Axis(
    f[1, 1],
    xlabel = "Time",
    ylabel = "Number"
    )

lines!(ax, sir_array[4, :], sir_array[1, :])
lines!(ax, sir_array[4, :], sir_array[2, :])
lines!(ax, sir_array[4, :], sir_array[3, :])

>>>>

```{julia}
using JumpProcesses, StochasticDiffEq, DifferentialEquations, LinearAlgebra
using Catalyst, Random, Statistics, DataFrames, CairoMakie, ProgressMeter

@@

```{julia}
X = 450
Y = 50
Z = 0
N = X + Y + Z
u₀ = [X, Y, Z, N]
```

```{julia}
R₀ = 12
γ = 1/8
μ = 1 / (365 * 62.5)
β = R₀ * (γ + μ)
p = (; β, γ, μ)
```

```{julia}
tmax = 100 * 365
δt = 0.001
tspan = (0.0, tmax)
tlength = length(tspan[1]:δt:tspan[2])
Random.seed!(1234);
```

```{julia}
rate1(u, p, t) = p.β * u[1] * u[2] / u[4]  # β*X*Y / N

function affect1!(integrator)
    integrator.u[1] -= 1         # X -> X - 1
    integrator.u[2] += 1         # Y -> Y + 1
    nothing
end

jump = ConstantRateJump(rate1, affect1!)
```

```{julia}
rate2(u, p, t) = p.γ * u[2]         # γ*Y

function affect2!(integrator)
    integrator.u[2] -= 1        # Y -> Y - 1
    integrator.u[3] += 1        # Z -> Z + 1
    nothing
end

jump2 = ConstantRateJump(rate2, affect2!)
```

```{julia}
birth_rate(u, p, t) = p.μ * u[4]    # μ * N

function birth_affect!(integrator)
    integrator.u[1] += 1
    integrator.u[4] += 1
    nothing
end

birth_jump = ConstantRateJump(birth_rate, birth_affect!)
```

```{julia}
death_rate(u, p, t) = p.μ * u[4]    # μ * N

function death_affect!(integrator)
    map!(
        i -> ifelse(integrator[i] > 0, integrator.u[i] - 1, integrator.u[i]),
        integrator.u,
        eachindex(integrator.u)
    )

death_jump = ConstantRateJump(death_rate, death_affect!)
```

```{julia}
prob = DiscreteProblem(u₀, tspan, p)
jump_prob = JumpProblem(prob, Direct(), jump, jump2, birth_jump, death_jump)
sol = solve(jump_prob, SSAStepper())

sol_array = Array(sol)
```
```{julia}
f = Figure()
ax = Axis(
    f[1, 1],
    xlabel = "Time",
    ylabel = "Number"
)

lines!(ax, sol.t, sol_array[1, :], linewidth = 3, label = "X(t)")
lines!(ax, sol.t, sol_array[2, :], linewidth = 3, label = "Y(t)")
lines!(ax, sol.t, sol_array[3, :], linewidth = 3, label = "Z(t)")
lines!(ax, sol.t, sol_array[4, :], linewidth = 3, label = "N(t)")

axislegend(ax)
